Optimal management of a stochastically varying population when policy adjustment is costly

### 2014-07-30 manuscripts/manuscript.Rmd

$$N_{t+1} = Z_t \frac{A (N_t - h_t)}{1 + B (N_t - h_t)}, \label{eq:state_equation}$$

@@

We compare this base case to three alternative problem formulations,
each reflecting different plausible functional forms that costs of
policy adjustment could take. In each, we assume managers can adjust the
quota set in the fishery $h_t$ in a given year and that any policy
adjustment costs are associated with changes to this control variable.

@@

Each of the policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal net
present value of fishery, broadly following the example of @Bovenberg2008.

@@

We focus our examination of the models on how the type and severity of
policy adjustment costs affects the optimal management strategy. To
simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the intrinsic rate of increase of
the fish population is 1 and the equilibrium biomass for the equivalent
deterministic model without harvesting is 10; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn. (1)). To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with 
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.

@@

For example, when small policy adjustment costs cost little but large
adjustments are expensive ($\Pi_2$), we see the smoothing signal that we
might have expected (see also @Ludwig1980 for this particular case). As
cost penalties increase in severity (i.e., $c_i$ increases), the
variance in quotas through time decreases and the autocorrelation of
quotas through time increases. This smoothing of the harvest quotas
through time has knock-on effects for the corresponding optimal stock
sizes, which also become more autocorrelated as the scaling on policy
adjustment costs ($c_i$) is increased. The autocorrelation in stock
sizes here reflects the lag involved in a more modest increase in
harvests taking longer to eliminate the signal of strong recruitment
events, when compared to a more immediately responsive constant
escapement policy.

@@

Finally, the case where policy adjustment costs scale linearly with the
size of the adjustment ($\Pi_1$ in Eqn. 4) appear to be something of a
middle of the road strategy, in that increasing the severity of policy
adjustment costs (increasing $c_i$ has little effect on the variance or
autocorrelation of optimal harvest rates or stock sizes. To reveal the
particular impact of policy adjustment costs of this type requires a
more targeted summary statistic. Specifically, for each run we
calculated the frequency with which the optimal policy involved
maintaining a positive quota across multiple time steps unaltered. This
type of policy is arguably the most commonly observed behavior in TAC
management, but is one that is almost never observed to be part of the
optimal management strategy in the basic model without policy adjustment
costs (Eqn. 3) or when optimizing against $\Pi_2$ or $\Pi_3$.

@@

Histograms for the different contributions labelled $[1]-[4]$ here when
taken across 100 replicate runs are shown are shown in Figure 5.
<!--(contribution $[1]$ red, $[2]$ green, $[3]$ blue, and $[4]$ mauve).--> The
cost of assuming policy adjustments are present when in fact they are
absent is given by subtracting the contribution to the first infinite
sum from term $[1]$ from the contribution to the second infinite sum of
term $[3]$. (Note that convergence properties
required to do this are met because we assume a positive discount rate).
This cost is positive because the optimal controls recommended when
assuming policy adjustment costs are present do not track variations in
stock size as closely. Therefore, this harvesting strategy provides
lower net revenue dockside.

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure 7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10%
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

@@

These patterns hold across the different size reductions shown (from 10%
to 25%), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.  Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.", dev=c("pdf", "svg", "png"), fig.width=7}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

@@

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect
on annual quotas and those that do not. The few past studies that
incorporate policy adjustment costs in models of fisheries management
[@Ludwig1980; @Feichtinger1994; @Wirl1999] assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (our $\Pi_2$ formulation in Eqn.) In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smoothes inter-annual
variation in the catch quotas that would be recommended. However, as
catches become less variable through time, the variability in the fish
population itself increases. In essence, the optimal policy no longer
soaks up as much of the natural variability in recruitment that is
affecting the population. Also, the catches and remaining stock sizes
corresponding to optimal management become more autocorrelated in time,
because it takes the fishery longer to harvest down peaks in abundance
that follow large recruitment pulses. The effects of smoothing here are
similar to those predicted when assuming the cost per unit effort is
increasing in the amount of effort expended (e.g., @Lewis1981, @McGough2009; 
see Supporting Information for a summary of the relevant
results) as opposed to associating extra costs with changes to policy
*per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

Standing in sharp contrast to these smoothing predictions is our
finding that including policy adjustment costs can actually _increase_ the
variability of quotas through time if there is a fixed cost associated
with making any changes to current policy (e.g. costs of running relevant
stakeholder meetings and public consultations on proposed policy changes,
$\Pi_3$). With this formulation, as stock sizes vary in response to
recruitment, the fishery manager must balance the cost of brokering
a policy change with the cost in forgone fisheries revenue from not
responding to favorable recruitment pulses. The optimal policy involves
ignoring small variations in recruitment, but then assigns a larger quota
when particularly strong recruitment years arise than would have been
the case in the absence of policy adjustment costs. The other functional
form we consider, in which the costs of policy adjustment scale linearly
with the size of the adjustment ($\Pi_1$), has less obvious effects on
optimal policies. But interestingly, it is the only formulation that
most frequently produces stretches of strictly positive but unchanging
quotas of the type most commonly encountered in real-world applications.

@@

As with any modeling analysis, our formulation makes many assumptions. For
example, we assume that the fishery in question is being optimally managed
by a policy-maker who acts as the 'sole owner' of the stock. This approach
is different to models that assume fisheries are not well-managed,
e.g. by assuming open access conditions (references [^1]),
or ones that derive policy recommendations endogenously by modeling
interactions between different stakeholders [@Kaitala1993, @Laukkanen2003].
As such, we anticipate that our approach will be more relevant to
some fisheries, particularly domestic fisheries in developed countries
that are subject to relatively strong regulatory regimes, than to others
(e.g., artisanal fisheries that are subject to weaker regulation). To
focus on the effects of policy adjustment costs, we focused narrowly
on the basic model specification of @Reed1979 and examined how the
predictions of this classic problem were changed by introducing policy
adjustment costs of different forms. However, there have been many
elaborations on Reed's basic approach that increase the realism of
the optimization models involved by relaxing other assumptions (see for
example @Sethi2005; @Singh2006; @McGough2009). For
example, in following Reed's approach closely another important
assumption we made is that the fishery manager observes the current stock
size correctly when setting the annual quota (see @Clark1986,
@Williams2001 for alternative specifications). However, predictions of
current cohort sizes from stock assessments are subject to uncertainty
[@Ralston2011]. Were uncertainty regarding reported recruitment
pulses also included, an optimizing manager might moderate their policy
responses to initial reports of particularly good recruitment years
increase yet further.

>>>>

\begin{equation}
N_{t+1} = Z_t \frac{A (N_t - h_t)}{1 + B (N_t - h_t)}, \label{eq:state_equation}
\end{equation}

@@

We compare this base case to three alternative problem formulations,
each reflecting different plausible functional forms that costs of
policy adjustment could take. In each, we assume managers can adjust the
quota set in the fishery $h_t$ in a given year and that any policy
adjustment costs are associated with changes to this control variable.
In each case, we assume there is no cost in initially setting the harvest
policy at time 0.

@@

Each policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal net
present value of fishery, broadly following the example of @Bovenberg2008.

@@

We focus our examination of the models on how the type and severity of
policy adjustment costs affects the optimal management strategy. To
simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the intrinsic rate of increase of
the fish population is 1 and the equilibrium biomass for the equivalent
deterministic model without harvesting is 10; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}. To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with 
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.

@@

For example, when small policy adjustments cost little but large
adjustments are expensive ($\Pi_2$), we see the smoothing signal that we
might have expected (see also @Ludwig1980 for this particular case). As
cost penalties increase in severity (i.e., $c_i$ increases), the
variance in quotas through time decreases and the autocorrelation of
quotas through time increases. This smoothing of the harvest quotas
through time has knock-on effects for the corresponding optimal stock
sizes, which also become more autocorrelated as the scaling on policy
adjustment costs ($c_i$) is increased. The autocorrelation in stock
sizes here reflects the lag involved in a more modest increase in
harvests taking longer to eliminate the signal of strong recruitment
events, when compared to a more immediately responsive constant
escapement policy.

@@

Finally, the case where policy adjustment costs scale linearly with the
size of the adjustment ($\Pi_1$) appear to be something of a
middle of the road strategy, in that increasing the severity of policy
adjustment costs (increasing $c_i$ has little effect on the variance or
autocorrelation of optimal harvest rates or stock sizes. To reveal the
particular impact of policy adjustment costs of this type requires a
more targeted summary statistic. Specifically, for each run we
calculated the frequency with which the optimal policy involved
maintaining a positive quota across multiple time steps unaltered. This
type of policy is arguably the most commonly observed behavior in TAC
management, but is one that is very rarely observed to be part of the
optimal management strategy in the basic model without policy adjustment
costs (Eqn. 3) or when optimizing against $\Pi_2$ or $\Pi_3$.

@@

Histograms for the different contributions labelled $[1]-[4]$ here when
taken across 100 replicate runs are shown in Figure 5. Observe that the red
distributions (dockside profits) are relatively unchanged between policies based on the adjustment-free assumption (Reed, bottom row) vs when adjustments are properly accounted for (adjustment penalty, top row), while the blue distributions (costs for adjustment) are much larger in the bottom row then
in the top (when these costs have not been accounted for vs when they have
been taken into account.)  The cost of assuming policy adjustments are present when in fact they are
absent is given by subtracting the contribution to the first infinite
sum from term $[1]$ from the contribution to the second infinite sum of
term $[3]$. (Note that convergence properties
required to do this are met because we assume a positive discount rate).
This cost is positive because the optimal controls recommended when
assuming policy adjustment costs are present do not track variations in
stock size as closely. Therefore, this harvesting strategy provides
lower net revenue dockside.

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure 7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

@@

These patterns hold across the different size reductions shown (from 10%
to 25%, panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.  Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% (panel A) of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.", dev=c("pdf", "svg", "png"), fig.width=7}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

@@

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect
on annual quotas and those that do not. The few past studies that
incorporate policy adjustment costs in models of fisheries management
[@Ludwig1980; @Feichtinger1994; @Wirl1999] assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (our $\Pi_2$ formulation) In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smoothes inter-annual
variation in the catch quotas that would be recommended. Also, the catches and remaining stock sizes
corresponding to optimal management become more autocorrelated in time,
because it takes the fishery longer to harvest down peaks in abundance
that follow large recruitment pulses. The effects of smoothing here are
similar to those predicted when assuming the cost per unit effort is
increasing in the amount of effort expended (e.g., @Lewis1981, @McGough2009; 
see Supporting Information for a summary of the relevant
results) as opposed to associating extra costs with changes to policy
*per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

Standing in sharp contrast to these smoothing predictions is our
finding that including policy adjustment costs can actually _increase_ the
variability of quotas through time if there is a fixed cost associated
with making any changes to current policy (e.g. costs of running relevant
stakeholder meetings and public consultations on proposed policy changes,
$\Pi_3$). With this formulation, as stock sizes vary in response to
recruitment, the fishery manager must balance the cost of brokering
a policy change with the cost in forgone fisheries revenue from not
responding to favorable recruitment pulses. The optimal policy involves
ignoring small variations in recruitment, but then assigns a larger quota
when particularly strong recruitment years arise than would have been
the case in the absence of policy adjustment costs. The other functional
form we consider, in which the costs of policy adjustment scale linearly
with the size of the adjustment ($\Pi_1$), has less obvious effects on
optimal policies. But interestingly, it is the formulation that
most frequently produces stretches of strictly positive but unchanging
quotas of the type most commonly encountered in real-world applications.

@@

As with any modeling analysis, our formulation makes many assumptions. For
example, we assume that the fishery in question is being optimally managed
by a policy-maker who acts as the 'sole owner' of the stock. This approach
is different to models that assume fisheries are not well-managed,
e.g. by assuming open access conditions (references [^1]),
or ones that derive policy recommendations endogenously by modeling
interactions between different stakeholders [@Kaitala1993; @Laukkanen2003].
As such, we anticipate that our approach will be more relevant to
some fisheries, particularly domestic fisheries in developed countries
that are subject to relatively strong regulatory regimes, than to others
(e.g., artisanal fisheries that are subject to weaker regulation). To
focus on the effects of policy adjustment costs, we focused narrowly
on the basic model specification of @Reed1979 and examined how the
predictions of this classic problem were changed by introducing policy
adjustment costs of different forms. However, there have been many
elaborations on Reed's basic approach that increase the realism of
the optimization models involved by relaxing other assumptions (see for
example @Sethi2005; @Singh2006; @McGough2009). For
example, in following Reed's approach closely another important
assumption we made is that the fishery manager observes the current stock
size correctly when setting the annual quota (see @Clark1986,
@Williams2001 for alternative specifications). However, predictions of
current cohort sizes from stock assessments are subject to uncertainty
[@Ralston2011]. Were uncertainty regarding reported recruitment
pulses also included, an optimizing manager might moderate their policy
responses to initial reports of particularly good recruitment years
increase yet further.
